Co-moving stars in \textsl{Gaia DR1}:\\ An abundance of very wide separation co-moving pairs

### 2017-02-03 paper/ms.tex
more work on comments

We construct an initial sample of star pairs to search for co-moving pairs as
follows.
We first apply a global parallax signal-to-noise cut, $\snr > 8$,  to the \tgas,
which leaves 619,618 stars.
Then, for each star we establish an initial sample of possible
co-moving partners by selecting all other stars that satisfy two criteria:
separation less than 10~pc and difference
in (point-estimate) tangential velocity less than $\absdvtan < 10$~\kms.
We ultimately build a statistical model that incorporates the covariances of the
data, but for these initial cuts and for visualizations we use a point-estimate
of the distance by applying a correction for the Lutz-Kelker bias
(\citealt{Lutz:1973aa}):
\begin{equation}
  \hat{r} = 1000 \, \left[\frac{\varpi}{2} \,
    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
    \label{eq:dist}
\end{equation}
where $\varpi$ is the parallax in mas.
An estimate for the difference in tangential velocity between two stars is,
then,
\begin{equation}
  \absdvtan = |\hat{r}_1 \vec\mu_1 - \hat{r}_2 \vec\mu_2|
\end{equation}
where $\vec\mu = (\mu_\alpha, \mu_\delta)$.

@@

The abundance of pairs of stars with small velocity difference in
\figname~\ref{fig:dv-sep} suggests that there are a
significant number of co-moving pairs in the \tgas\ data at a range
of separations.
Here, we develop a method to select high-confidence co-moving
pairs that properly incorporates the uncertainties associated with the
\gaia\ data. We make the following assumptions in order to construct a
statistical model (a likelihood function with explicit priors on our
parameters):
\begin{itemize}
  \item We assume that the uncertainties in the data---parallax, $\varpi$, and
    two proper motion components, $\propm = (\begin{array}[t]{c c} \mu_\alpha &
    \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances
    $\mat{C}$. The values of covariances are provided as part of the \gaia\ data
    \citep{Lindegren:2012aa,Lindegren:2016aa}.
  \item We assume that the 3D velocities of stars in a given pair
    $(\vec{v}_i, \vec{v}_j)$ (relative to the solar system
    barycenter) are either (1) the same with a small (Gaussian) dispersion $s$
    or (2) independent.
    In both cases, velocity is drawn from
    the velocity prior $p(\vec{v})$.
\end{itemize}

>>>>

We construct an initial sample of star pairs to search for co-moving pairs as
follows.
We first apply a global parallax signal-to-noise cut, $\snr > 8$,  to the \tgas,
which leaves 619,618 stars.
Then, for each star we establish an initial sample of possible
co-moving partners by selecting all other stars that satisfy two criteria:
separation less than 10~pc and difference
in (point-estimate) tangential velocity less than $\absdvtan < 10$~\kms.
We ultimately build a statistical model that incorporates the covariances of the
data, but for these initial cuts and for visualizations we use a point-estimate
of the distance by applying a correction for the Lutz-Kelker bias
(\citealt{Lutz:1973aa}):
\begin{equation}
  \hat{r} = 1000 \, \left[\frac{\varpi}{2} \,
    \left(1 + \sqrt{1 - \frac{16}{\snr^2}} \right) \right]^{-1} \, {\rm pc}
    \label{eq:dist}
\end{equation}
where $\varpi$ is the parallax in mas.
An estimate for the difference in tangential velocity between two stars is,
then,
\begin{equation}
  \absdvtan = |\hat{r}_1 \vec\mu_1 - \hat{r}_2 \vec\mu_2|
\end{equation}
where $\vec\mu = (\mu_{\alpha^*}, \mu_\delta)$.
\footnote{$\mu_{\alpha^*}$ is proper motion
in right ascension direction, $\mu_{\alpha^*} = \mu_\alpha \cos\delta$}

@@

The abundance of pairs of stars with small velocity difference in
\figname~\ref{fig:dv-sep} suggests that there are a
significant number of co-moving pairs in the \tgas\ data at a range
of separations.
Here, we develop a method to select high-confidence co-moving
pairs that properly incorporates the uncertainties associated with the
\gaia\ data. We make the following assumptions in order to construct a
statistical model (a likelihood function with explicit priors on our
parameters):
\begin{itemize}
  \item We assume that the uncertainties in the data---parallax, $\varpi$, and
    two proper motion components, $\propm = (\begin{array}[t]{c c} \mu_{\alpha^*} &
    \mu_\delta\end{array})^\mathsf{T}$---are Gaussian with known covariances
    $\mat{C}$. The values of covariances are provided as part of the \gaia\ data
    \citep{Lindegren:2012aa,Lindegren:2016aa}.
  \item We assume that the 3D velocities of stars in a given pair
    $(\vec{v}_i, \vec{v}_j)$ (relative to the solar system
    barycenter) are either (1) the same with a small (Gaussian) dispersion $s$
    or (2) independent.
    In both cases, velocity is drawn from
    the velocity prior $p(\vec{v})$.
\end{itemize}
